A Machine Learning Approach to MLB Catcher Framing

### 2021-12-10 Slides.Rmd
Finish slides

```{r setup, include=FALSE}
library(tidyverse)
library(glmnet)
library(gbm)
library(randomForest)
knitr::opts_chunk$set(echo = FALSE)
load("Data/usable_data.Rda")
```

@@

- We used:
- Pitch type and pitch release speed, position, and spin rate
- Whether or not the pitcher and batter are right or left handed
- Count, number of outs during the at-bat, and inning number
- Where the pitch landed
- Whether the game was played home or away
- How tall the batter is

@@

ggplot(df_gbm_final, aes(x = interaction_depth, y = accuracy, color = as.character(lambda))) +
  geom_line() +
  theme_minimal()+
  geom_point() +
  geom_point(aes(x = 16, y = 0.935), color = "red") +
  labs(x = "Accuracy", y = "Interactin Depth",
       color = "Shrinkage",
       title = "Boosting Accuracy by Interaction Depth and Shrinkage")
```

@@

```{r}
df_rf %>%
  rename(parameter = nodesize) %>%
  bind_rows(final_df_gbm %>%
              group_by(trees) %>%
              summarize(accuracy = mean(accuracy)) %>%
              arrange(desc(accuracy)) %>%
              mutate(model = "boosting") %>%
              select(model, accuracy, trees) %>%
              rename(parameter = trees)) %>%
  arrange(desc(accuracy))
```

@@

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(glmnet)
library(gbm)
library(randomForest)
library(caret)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
load("Data/usable_data.Rda")
```

@@

- We used:
- Pitch type, pitch release speed, position, and spin rate
- Whether or not the pitcher and batter are right or left handed
- Count, number of outs during the at-bat, and inning number
- Where the pitch landed
- Whether the game was played home or away
- How tall the batter is

@@

ggplot(df_gbm_final, aes(x = interaction_depth, y = accuracy, color = as.character(lambda))) +
  geom_line() +
  theme_minimal()+
  geom_point() +
  geom_point(aes(x = 16, y = 0.935), color = "red") +
  labs(y = "Accuracy", x = "Interactin Depth",
       color = "Shrinkage",
       title = "Boosting Accuracy by Interaction Depth and Shrinkage")
```

@@

```{r}
df_rf %>%
  rename(parameter = nodesize) %>%
  bind_rows(final_df_gbm %>%
              group_by(trees) %>%
              summarize(accuracy = mean(accuracy)) %>%
              arrange(desc(accuracy)) %>%
              mutate(model = "boosting") %>%
              select(model, accuracy, trees) %>%
              rename(parameter = trees)) %>%
  arrange(desc(accuracy)) %>%
  kable
```

## Variable Importance

```{r, message=FALSE,echo=FALSE, cache = T, warning=FALSE}
load("Final-Model/final_boosting_model.Rda")
var = summary.gbm(final_model, plotit = F)
ggplot(var, aes(x = fct_reorder(var, rel.inf), y = rel.inf, fill = rel.inf)) +
  geom_col() +
  coord_flip() +
  labs(y = "Relative Influence",
       title = "Predictor Variable Importance") +
  theme_minimal()+
  theme(axis.title.y = element_blank()) +
  guides(fill = F)
```

@@

```{r, fig.width=2, fig.height=2}
# Analysis examples
ggplot()+
  geom_path(data = sz, aes(x = x, y = y)) +
  xlim(-1.8, 1.8) +
  ylim(0.5, 5) +
  coord_equal() +
  geom_point(aes(x = .5, y = 2.5)) +
  geom_text(aes(x = 0.6, y = 2.8, label = "0.9")) +
  geom_point(aes(x = -.5, y = 4)) +
  geom_text(aes(x = -0.4, y = 4.3, label = "0.2")) +
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()) 
```

- If strike: 1 - 0.9 = 0.1 | 1 - 0.2 = 0.8
- If ball: 0.9 * -1 = -0.9 | 0.2 * -1 = -0.2

\tiny
```{r, cache= T, message=FALSE}
load("Data/pitchData.Rda")
# Analysis
data = data %>%
  mutate(strike_prob = predict(final_model, data, type = "response"),
         strike = as.numeric(as.character(strike)))

data = data %>%
  mutate(credit = case_when(
    strike == 1 ~ 1 - strike_prob,
    strike == 0 ~ strike_prob * -1))
```

```{r}

# Leader board
data %>%
  group_by(catcher_name) %>%
  summarize(strikes_above_avg = sum(credit),
            pitches_caught = n(),
            strikes_aa_100 = strikes_above_avg / (pitches_caught / 100)) %>%
  filter(pitches_caught > 2000) %>%
  arrange(desc(strikes_above_avg)) %>% head(10) %>%
  rename("Catcher Name" = catcher_name,
         "Strikes Above Average" = strikes_above_avg,
         "Pitches Caught" = pitches_caught,
         "Strikes Above Average Per 100 Pitches Caught" = strikes_aa_100) %>% 
  kable

data %>%
  group_by(catcher_name) %>%
  summarize(strikes_above_avg = sum(credit),
            pitches_caught = n(),
            strikes_aa_100 = strikes_above_avg / (pitches_caught / 100)) %>%
  filter(pitches_caught > 2000) %>%
  arrange(strikes_above_avg) %>% head(10) %>%
  rename("Catcher Name" = catcher_name,
         "Strikes Above Average" = strikes_above_avg,
         "Pitches Caught" = pitches_caught,
         "Strikes Above Average Per 100 Pitches Caught" = strikes_aa_100) %>% 
  kable
```

## Video Example 1

- Strike (0.00108 strike probability)
  - https://baseballsavant.mlb.com/sporty-videos?playId=2a91c3cf-e5d9-41f9-86e4-88700536a6de
- Austin Nola gained his team 1 - 0.00108 = 0.9982 strikes from this pitch

## Video Example 2

- Ball (0.99087 strike probability)
  - https://baseballsavant.mlb.com/sporty-videos?playId=3509cf12-7e2c-4beb-88af-d3b3cc736c02
- Kevan Smith lost his team -0.99087 strikes from this pitch
